
---
title: "Choosing a Subset"
output: 
  html_document:
    fig_caption: no
    number_sections: no
    toc: no
    toc_float: false
    collapsed: no.
---

```{r set-options, echo=FALSE}
options(width = 105)
knitr::opts_chunk$set(dev='png', dpi=300, cache=TRUE)
pdf.options(useDingbats = TRUE)
```

```{r}
# Set CRAN mirror directly
options(repos = "https://cloud.r-project.org")

# Now try installing the package
install.packages("plyr")

library("maps")
library("mapdata")
library("ggplot2")
library("readr")
library(plyr)
```

```{r}
soil_data <- read.csv(file='./SRDB_V5_1827/data/srdb-data-V5.csv')
names(soil_data)
attach(soil_data)
```
Lets take a look at where the sample are from:
```{r}
background_map <- map_data("worldHires")
soil_samps <- read_csv("SRDB_V5_1827/data/srdb-data-V5.csv")

```
```{r}
#adding our map:
ggplot() +
  geom_polygon(data = background_map, aes(x = long, y = lat, group = group), fill = "lightgrey") +
  geom_point(data = soil_samps, aes(x = Longitude, y = Latitude), color = "red", size = 0.5) +
  coord_fixed() +
  labs(title = "Map of Soil Sample Sites", x = "Longitude", y = "Latitude")
```


```{r}
#getting the frequency of different countries:
country_counts <- count(soil_data, 'Country')
head(country_counts)
```
```{r}
#lets sort them to see the most frequent countries:
attach(country_counts)
country_counts_sorted <- country_counts[order(-freq),]
head(country_counts_sorted)
```
We can visualize this with a barplot:

```{r}
#create barplot:
top_5 <- country_counts_sorted[1:5,]
y_top_5 <- top_5[['freq']]
x_top_5 <- top_5[['Country']]
barplot(y_top_5, names.arg=x_top_5, horiz=TRUE, xlab = 'Frequency', main = 'Top 5 Countries Represented', las = 1, col = 'darkgreen')
```
We see here that China and the USA make up a very large part of this data set. I want to see which of the two has the lowest percentage of missing values, then I will choose that subset of the data to work on:

## Subsetting China and USA:

First we will only grab Countries China and USA:
```{r}
subset_list <- c("China", "USA")
china_usa = soil_data[soil_data$Country %in% subset_list, ]
```
And we can now see which has a lower percentage of missing values, we will also look at the distribution of quality flags for each group, because this will give us a better idea about the nature of information contained there:

```{r}
china <- china_usa[china_usa$Country == "China",]
percent_missing_china <- mean(is.na(china)) * 100
percent_missing_china
```
Lets now look at the percent missing from USA:
```{r}
usa <- china_usa[china_usa$Country == "USA",]
percent_mising_usa <- mean(is.na(usa)) * 100
percent_mising_usa
```
Great we can see that china has a lower percentage of missing data. 

#Inspection of Quality Flags:
For reference we can describe each quality flag as follows: \
Q0 default/none\
Q01 estimated from figure\
Q02 data from another study\
Q03 data estimated--other\
Q04 potentially useful future data\
Q10 potential problem with data\
Q11 suspected problem with data\
Q12 known problem with data\
Q13 duplicate\
Q14 inconsistency\


Lets take a look at the distribution of flags for each group:
```{r}
usa_qf_counts <- count(usa, "Quality_flag")
china_qf_counts <- count(china, "Quality_flag")
```
```{r}
usa_qf_counts
```
```{r}
china_qf_counts
```

Alright we can pretty easily see that China has more data, and lest potental problems with the quality of each observation. 

### Moving Forward with the China Subset:

The next step is cleaning the data, addressing missing values, and selecting our features of interest. 

The first thing we will do is select data in China with Quality flags = Q0 (or none), Q1, Q2, and Q3. I feel it would be potentially misleading or problematic to keep observations with the other QF's, we will then save as a data set and move on the the cleaning and feature selection section!
```{r}
ok_flags = c(NULL, "Q01", "Q1", "Q02", "Q03")
china_soil_respiration <- soil_data[(soil_data$Country == "China") & (soil_data$Quality_flag %in% ok_flags), ]
head(china_soil_respiration)
```
Now lets export this and move on to the next section: Data Cleaning and Feature Selection
```{r}
# Save the dataframe as a CSV file
write.csv(china_soil_respiration, file = "china_soil_respiration.csv", row.names = FALSE)
```



